---
title: "Sleep and depression data"
author: "Ana"
date: "11/02/2021"
output: html_document
---


```{r message = FALSE, include=FALSE}
library(tidyverse)
source("../helper_functions.R")
library(RColorBrewer)
```

#Tabela sleep depression
```{r echo=F}
sleep_depression <- readRDS("../Data/Processed/df_sleep_depression")

sleep_depression <- sleep_depression %>% 
  filter(PRIMDIAG %in% c("Idiopathic Parkinson's Disease", "Healthy"))

sleep_depression
```

# Depression rate in patients with idiopathic PD

> Does PD patients present a higher rates of depression vs. healthy patients? 
>
> -Yes:   Depression: PD = 17,98% x Healthy = 4,8%

```{r echo = FALSE}
healthy_vs_depressive <- sleep_depression %>%
  select(PATNO, year, depressive_status, depression_category, PRIMDIAG) %>%
  mutate(PRIMDIAG = factor(x = PRIMDIAG,
                           levels = c("Idiopathic Parkinson's Disease",
                                      "Healthy"))) %>%
  mutate(depression_category = factor(x = depression_category,
                                      levels = c("Severe depression", 
                                                 "Moderate depression",
                                                 "Mild depression",
                                                 "Normal"))) %>%
  count(depression_category,PRIMDIAG) %>%
  group_by(PRIMDIAG) %>%
  mutate(perc_pd_depression = n/sum(n) *100) 

healthy_vs_depressive %>% 
  ggplot(aes(x = depression_category, 
             y = perc_pd_depression, 
             fill = PRIMDIAG)) +
  geom_col(position = position_dodge()) +
  geom_label(aes(label = round(perc_pd_depression, 2) %>%  paste0(. , "%")),
             position = position_dodge(width = 0.9),
             hjust = 0,
             fontface = "bold",
             label.size = 0,
             alpha = 0,
             size = 2.5,
             show.legend = FALSE) +
  scale_fill_manual(values = safe_colorblind_palette) +
  scale_x_discrete(labels = c("Severe depression", "Moderate depression", "Mild depression", "Non depressive")) +
  theme_classic() +
  labs(title = str_wrap(string = "Depression rate in Parkinson's disease patients",
                        width = 60),
       subtitle = str_wrap(string = "Higher percentage of depression in Parkinson's disease patients",
                           width = 40),
       x = "Depression Severity",
       y = "%",
       fill = "") +
  ylim(0,100) +
  theme(plot.title = element_text(hjust = 0.5, 
                                  size = 12,
                                  face = "bold",
                                  margin = margin(0,0,5,0)),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 11,
                                  margin = margin(0,0,30,0)),
        axis.text = element_text(size = 10),
        axis.title.y = element_text(size = 10,
                                    margin = margin(0,15,0,0)),
        axis.title.x = element_text(size = 10,
                                    margin = margin(10,0,0,0)),
        legend.position = "right",
        legend.justification = "center",
        legend.margin = margin(0,0,0,30)) +
  coord_flip()
```

# Depressive status in PD patients

> PD is associated with lower percentage of patients in the constant healthy groups compared with healthy patients
>
> Higher percentage of depression in PD patients is associated with:
>
> -   higher percentage of PD patients whom became depressive during the analysis, compared with healthy patients
>
> -   higher percentage of PD whom became healthy - "cured", compared with healthy patients
>
> -   higher levels of constant depression throughout the analysis, compared with healthy patients (although the percentage is lower compared with the other categories from the depression spectrum)

```{r echo = FALSE}
sleep_depression %>% 
  select(PATNO, year, depressive_status, depression_category, PRIMDIAG) %>% 
  filter(depressive_status != "single_data") %>% 
  # mutate(PRIMDIAG = factor(x = PRIMDIAG,
  #                          levels = c("17",
  #                                     "01"))) %>% 
  mutate(depressive_status = factor(x = depressive_status,
                                    levels = c("constant_healthy",
                                               "constant_depressive",
                                               "became_healthy",
                                               "became_depressive"))) %>%
  count(depressive_status,PRIMDIAG) %>% 
  group_by(PRIMDIAG) %>% 
  mutate(perc_pd_depression_status = n/sum(n) *100) %>% 
  ggplot(aes(x = depressive_status,
             y = perc_pd_depression_status,
             fill = PRIMDIAG)) +
  geom_col(position = position_dodge()) +
  geom_label(aes(label = round(perc_pd_depression_status, 2) %>% paste0(. , "%")),
             position = position_dodge(width = 0.9), 
              hjust = 0,
              fontface = "bold",
              label.size = 0,
              alpha = 0,
              size = 2.5,
             show.legend = FALSE) +
  scale_fill_manual(values = safe_colorblind_palette) +
  scale_x_discrete(labels = c("Constant healthy",
                              "Constant depressive",
                              "Became healthy",
                              "Became depressive")) +
  labs(title = "Depressive timeline in Parkinson's disease patients",
       x = "",
       y = "%",
       fill = "") +
  ylim(0,100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, 
                                  margin = margin(0,0,30,0)),
        axis.text = element_text(colour = "black",
                                 size = 10),
        axis.title = element_text(face = "bold",
                                  size = 12)) +
  coord_flip()
```

# Daytime sleepiness classification in depressive patients

```{r echo = FALSE}
sleep_depression %>% 
  select(PATNO, depressive, category_2) %>%
  count(depressive, category_2) %>%
  group_by(depressive) %>% 
  mutate(perc_depressive_sleep = n/sum(n) * 100) %>% 
  ggplot(aes(x = category_2,
             y = perc_depressive_sleep,
             fill = depressive)) +
  geom_col(position = position_dodge()) +
  geom_label(aes(label = round(perc_depressive_sleep, 2) %>% paste0(. , "%")),
             position = position_dodge(width = 0.9),
             hjust = 0,
             fontface = "bold",
             label.size = 0,
             alpha = 0,
             size = 2.5,
             show.legend = FALSE) +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_brewer(palette = "Dark2",
                    labels = c("Non depressive",
                               "Depressive")) +
  scale_x_discrete(labels = c("Lower Normal Daytime Sleepiness", 
                              "Higher Normal Daytime Sleepiness",
                              "Mild Excessive Daytime Sleepiness",
                              "Moderate Excessive Daytime Sleepiness",
                              "Severe Excessive Daytime Sleepiness")) +
  labs(title = "Daytime sleepiness in depressive patients",
       subtitle = str_wrap(string = "Higher excessive daytime slepiness in depressive patients",
                           width = 40),
       x = "",
       y = "%",
       fill = "") +
  theme_classic() +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  margin = margin(0,0,10,0),
                                  size = 12,
                                  face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, 
                                     margin = margin(0,0,20,0),
                                     size = 11),
        axis.text = element_text(colour = "black",
                                 size = 10),
        axis.title = element_text(size = 10)) +
  coord_flip()
 
```

# Daytime sleepiness classification in constant depressive patients

```{r echo = FALSE}
sleep_depression %>% 
  filter(depressive_status == c("constant_healthy", "constant_depressive")) %>%
  mutate(depressive_status = factor(x = depressive_status,
                                    levels = c("constant_healthy",
                                               "constant_depressive"))) %>% 
  count(depressive_status, category_2) %>% 
  group_by(depressive_status) %>% 
  mutate(percentage_ess_depressive = n/sum(n) *100) %>% 
   ggplot(aes(x = category_2, 
              y = percentage_ess_depressive,
              fill = depressive_status)) +
  geom_col(position = position_dodge()) +
  geom_label(aes(label = round(percentage_ess_depressive, 2) %>% paste0(. , "%")),
             position = position_dodge(width = 0.9),
             hjust = 0,
             fontface = "bold",
             label.size = 0,
             alpha = 0,
             size = 3,
             show.legend = FALSE) +
  scale_y_continuous(limits = c(0,100)) +
  scale_x_discrete(labels = c("Lower Normal Daytime Sleepiness", 
                              "Higher Normal Daytime Sleepiness",
                              "Mild Excessive Daytime Sleepiness",
                              "Moderate Excessive Daytime Sleepiness",
                              "Severe Excessive Daytime Sleepiness")) +
  scale_fill_brewer(palette = "Dark2",
                    labels = c("Constant healthy",
                             "Constante depressive")) +
 labs(title = str_wrap(string = "Daytime sleepiness classification in constant depressive patients",
                       width = 35),
     subtitle = str_wrap(string = "Patients with depression during all the analysis showed higher rates of excessive daytime sleepiness",
                         width = 60),
       x = "",
       y = "%",
       fill = "") +
  theme_classic() +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  margin = margin(0,0,10,0),
                                  size = 15,
                                  face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, 
                                  margin = margin(0,0,30,0),
                                  size = 12),
        axis.text = element_text(colour = "black",
                                 size = 10),
        axis.title = element_text(face = "bold",
                                  size = 10)) +
  coord_flip()
```
```

